\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes

### 2017-09-27 paper/paper.tex
Add separate supplementary section with plots
- change figure 6 to corresponding Drive-models of figure 5

- fix labeling clashes

\begin{figure}[H]
  %\includegraphics[width=.9\textwidth]{data-sum.pdf}
  \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.  Each number labels a unique drive brand model.   Axes are on the log scale.}
  \label{fig1}
\end{figure}

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g}, \pi_{g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{f(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-F(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L}$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.  All times are again in hours.

@@

Results showing estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:1}.

@@

\label{table:1}

@@

% \begin{center}
% {\large\bf SUPPLEMENTARY MATERIAL}
% \end{center}
% 
% \begin{description}
% 
% \item Put R Stan code here
% 
% \end{description}
%
%\section{Stan code for Model 4}
%\label{sec:stan-code}
%{\scriptsize
%\begin{verbatim}
%functions {
  %real sev_logpdf(real y, real mu, real sigma){
    %real z;
    %z = (y - mu) / sigma;
    %return -log(sigma) + z - exp(z);
  %}
  %
  %real sev_logccdf(real y, real mu, real sigma){
    %return -exp((y - mu) / sigma);
  %}
  %
  %real sev_logcdf(real y, real mu, real sigma){
    %return log1m_exp(-exp((y - mu) / sigma));
  %}
%}
%
%data {
  %int M;
  %int N_obs;
  %int N_cens;
  %real endtime_obs[N_obs];
  %real endtime_cens[N_cens];
  %real starttime_obs[N_obs];
  %real starttime_cens[N_cens];
  %int<lower=1, upper=M> dm_obs[N_obs];
  %int<lower=1, upper=M> dm_cens[N_cens];
  %vector<lower=0, upper=1>[2] p; # quantiles to model
%}
%transformed data{
  %vector[2] z_corr;
  %for(i in 1:2)
    %z_corr[i] = log(-1.0 * log1m(p[i])); # used to get location(mu) from quantile(t_p)
%}
%parameters{
  %real log_tp1;
  %real<lower=0> sigma1;
  %real eta_tp2;
  %real eta_s2;
  %real eta_pi;
  %real<lower=0> tau_tp2;
  %real<lower=0> tau_s2;
  %real<lower=0> tau_pi;
  %vector[M] log_tp2_raw;
  %real<lower=0, upper=1> sigma2[M];
  %vector[M] logit_pi_raw;
%}
%
%transformed parameters{
  %real mu1;
  %vector[M] mu2;
  %vector[M] log_pi;
  %mu1 = log_tp1 - sigma1 * z_corr[1];
  %for(m in 1:M){
    %mu2[m] = (eta_tp2 + tau_tp2 * log_tp2_raw[m]) - (sigma2[m] * z_corr[2]);
  %}
  %for(m in 1:M)
    %log_pi[m] = log_inv_logit(eta_pi + tau_pi * logit_pi_raw[m]);
%}
%
%model{
  %real tmp[2];
  %int m;
  %real logpi;
  %real mu_1;
  %real mu_2;
  %real sig_1;
  %real sig_2;
  %
  %for(i in 1:N_obs){
    %m = dm_obs[i];
    %logpi = log_pi[m];
    %mu_2 = mu2[m];
    %sig_2 = sigma2[m];
    %// numerator:   log( p * f1 * (1 - F2) + f2 * (1 - p * F1) )
    %//            = log( exp(log(p) + log(f1) + log(1 - F2)) + exp(log(f2) + log(1 - exp(log(p) + log(F1)))) )
    %tmp[1] = log_sum_exp(logpi + sev_logpdf(endtime_obs[i], mu1, sigma1) +
               %sev_logccdf(endtime_obs[i], mu_2, sig_2),
               %sev_logpdf( endtime_obs[i], mu_2, sig_2) + 
               %log1m_exp(logpi + sev_logcdf(endtime_obs[i], mu1, sigma1)
               %)
             %);
    %// denominator:  log((1 - p * F1) * (1 - F2))
    %//            =  log(1 - p * F1) + log(1 - F2)
    %tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_obs[i], mu1, sigma1)) + 
             %sev_logccdf(starttime_obs[i], mu_2, sig_2);
             %
    %target += tmp[1] - tmp[2];
  %}
  %
  %for(i in 1:N_cens){
    %m = dm_cens[i];
    %logpi = log_pi[m];
    %mu_2 = mu2[m];
    %sig_2 = sigma2[m];
  %
    %// numerator:   log((1 - p * F1) * (1 - F2))
    %//            =  log(1 - p * F1) + log(1 - F2)
    %tmp[1] = log1m_exp(logpi + sev_logcdf(endtime_cens[i], mu1, sigma1)) + 
             %sev_logccdf(endtime_cens[i], mu_2, sig_2);
    %// denominator:  log((1 - p * F1) * (1 - F2))
    %//            =  log(1 - p * F1) + log(1 - F2)
    %tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_cens[i], mu1, sigma1)) + 
             %sev_logccdf(starttime_cens[i], mu_2, sig_2);
             %
    %target += tmp[1] - tmp[2];
  %}
  %
  %log_tp1        ~ normal(7, 2);
  %log_tp2_raw    ~ normal(0, 1);
  %sigma1         ~ lognormal(0, 1);
  %sigma2         ~ lognormal(eta_s2, tau_s2);
  %logit_pi_raw   ~ normal(0, 1);
  %eta_tp2        ~ normal(9, 2);
  %eta_s2         ~ normal(0, 2);
  %eta_pi         ~ normal(-3, 1);
  %tau_tp2        ~ cauchy(0,1);
  %tau_s2         ~ cauchy(0,1);
  %tau_pi         ~ cauchy(0,1);
%}
%
%\end{verbatim}
%}
%
%\pagebreak
%\section{Posterior predictive graphical checks}
%\begin{figure}[H]
%\includegraphics[width=\textwidth]{ppcheck-1.pdf}
%\end{figure}
%\begin{figure}[H]
%\includegraphics[width=\textwidth]{ppcheck-2}
%\end{figure}
%\begin{figure}[H]
%\includegraphics[width=\textwidth]{ppcheck-3.pdf}
%\end{figure}
%\begin{figure}[H]
%\includegraphics[width=\textwidth]{ppcheck-4.pdf}
%\end{figure}
%\begin{figure}[H]
%\includegraphics[width=\textwidth]{ppcheck-5.pdf}
%\end{figure}
%\begin{figure}[H]
%\includegraphics[width=\textwidth]{ppcheck-6.pdf}
%\end{figure}
%\clearpage
%\bibliographystyle{plainnat}
% \bibliographystyle{plain}

>>>>

\begin{figure}[H]
  %\includegraphics[width=.9\textwidth]{data-sum.pdf}
  \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.  Each number labels a unique drive brand model.   Axes are on the log scale.}
  \label{drive-scatter}
\end{figure}

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g}, \pi_{g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{f(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-F(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.  All times are again in hours.

@@

Results showing estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:2}.

@@

\label{table:2}

@@
